---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dbpkgr

<!-- badges: start -->
<!-- badges: end -->

dbpkgr seeks to turn a database connection into a temporary R package to provide a set of convenience functions. Each created package supplies the following functions for the given connection:

- `_structure()` retrieves data.frame of catolog, schema and table names from information_schema.tables.
- `_list_schemas()` retrieves a character vector of schema names available.
- `_list_tables()` retrieves a character vector of table names available.
- `_query()` takes a SQL string and retrieves the result as a data.frame.
- `_execute()` takes an SQL string and executes it against the connection.

As well as a function for each and every table in the database, so you can take
advantage of autocomplete for table names.

The package is most functional where an `INFORMATION_SCHEMA` is provided by the
connection database management system (DBMS), which it can used to get the
`_structure()` and build queries like `FROM table_schema.table_name`. 

## Installation

You can install the development version of dbpkgr from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("davidfoord1/dbpkgr")
```

## Example

```{r example}
library(dbpkgr)
```

A demo connection with a database in memory:
```{r cars}
connection <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
DBI::dbWriteTable(connection, "mtcars", mtcars)
DBI::dbWriteTable(connection, "iris", iris)
```

Set up the package with a call to `dbp_package`, supplying the connection and
package name. Once the package is created you don't need to handle the
connection object anymore.

```{r package}
dbp_package(connection, "mydb")
```

The name will be used as a prefix for auto-complete friendly functions, with which you can get information from the connection:

```{r list_tables}
mydb_list_tables()
```

Tables get their own functions:

```{r}
mydb_mtcars()
```

They are lazy loaded - as a connection object that works with a dplyr/dbplyr chain:

```{r dplyr}
# median petal length/width ratio by species
mydb_iris() |>
  mutate(petal_ratio = Petal.Length / Petal.Width) |>
  summarise(petal_ratio = median(petal_ratio, na.rm = TRUE), .by = Species) |>
  arrange(desc(petal_ratio))
```

You can retrieve raw SQL query results:

```{r query}
mydb_query("SELECT cyl, COUNT(cyl) AS count FROM mtcars GROUP BY cyl")
```

And execute SQL statements that don't return a table:

```{r execute}
mydb_execute("CREATE TABLE table_1 (ID int PRIMARY KEY)")
mydb_list_tables()
mydb_execute("DROP TABLE table_1")
mydb_list_tables()
```

Basic documentation is automatically generated for the temporary package e.g. you could use 
```{r eval = FALSE}
help(mydb_structure)
```

